---
title: "Hands-on Exercise 1"
#subtitle: "Geospatial Data Wrangling with R"
description: "Geospatial Data Wrangling with R"
execute: 
  warning: false
# date: "`r Sys.Date()`"
date: 01/07/2024
date-format: "D MMM, YYYY" 
author: 
  - Wan Shen
sidebar: false
title-block-banner: true
categories:
  - Hands-On Exercise
title-block-categories: false
---

# 1.0 Getting Started

## 1.1 Installing and loading R packages

In this section, I will install and load tidyverse and sf packages.

```{r}
pacman::p_load(tidyverse, sf)
```

# 2.0 Importing a geospatial data in shapefile format

```{r}
#| echo: true #false if you do not want code display
#| eval: true #false if you do not want output
mpsz <- st_read(dsn = "../../data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")

cyclingpath = st_read(dsn = "../../data/geospatial", 
                         layer = "CyclingPathGazette")

preschool = st_read("../../data/geospatial/PreSchoolsLocation.kml")
```

# 3.0 Checking the Content of A Simple Feature Data Frame

```{r}
st_geometry(mpsz)
glimpse(mpsz)
head(mpsz, n=5)  
```

# 4.0 Plotting the Geospatial Data

```{r}
plot(mpsz)
plot(st_geometry(mpsz))
plot(mpsz["PLN_AREA_N"])
```

# 5.0 Working with Projection

## 5.1 Assigning EPSG code to a simple feature data frame

```{r}
st_crs(mpsz)
mpsz3414 <- st_set_crs(mpsz, 3414)
st_crs(mpsz3414)
```

## 5.2 Transforming the projection of preschool from wgs84 to svy21

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

# 6.0 Importing and Converting An Aspatial Data

## 6.1 Importing the aspatial data

```{r}
listings <- read_csv("../../data/aspatial/listings.csv")
list(listings) 
```

## 6.2 Creating a simple feature data frame from an aspatial data frame

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>% 
  st_transform(crs = 3414)
glimpse(listings_sf)
```

# 7.0 Geoprocessing with sf package

## 7.1 Buffering

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
buffer_cycling$AREA <- st_area(buffer_cycling)
sum(buffer_cycling$AREA)
```

## 7.2 Point-in-polygon count

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
summary(mpsz3414$`PreSch Count`)
top_n(mpsz3414, 1, `PreSch Count`)
mpsz3414$Area <- mpsz3414 %>% st_area()
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
mpsz3414['PreSch Density']
```

# 8.0 Exploratory Data Analysis (EDA)

```{r}
hist(mpsz3414$`PreSch Density`)

ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")

ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
