---
title: "Hands-on Exercise 8"
description: "Geographically Weighted Predictive Models"
execute: 
  warning: false
  freeze: true
# date: "`r Sys.Date()`"
date: 03/12/2024
date-format: "D MMM, YYYY" 
author: 
  - Wan Shen
sidebar: false
title-block-banner: true
categories:
  - Hands-On Exercise
title-block-categories: false
---

# 1.0 Getting Started

This code chunk performs 3 tasks:

-   A list called packages will be created and will consists of all the R packages required to accomplish this exercise.

-   Check if R packages on package have been installed in R and if not, they will be installed.

-   After all the R packages have been installed, they will be loaded.

```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse)
```

# 2.0 Preparing Data

## 2.1 Reading data file to rds

Reading the input data sets. It is in simple feature data frame.

```{r}
mdata <- read_rds("../../data/aspatial/mdata.rds")
```

## 2.2 Data Sampling

The entire data are split into training and test data sets with 65% and 35% respectively by using *initial_split()* of **rsample** package. rsample is one of the package of tigymodels.

```{r}
set.seed(1234)
resale_split <- initial_split(mdata, 
                              prop = 6.5/10,)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

```{r}
#| eval: false
write_rds(train_data, "../../data/model/train_data.rds")
write_rds(test_data, "../../data/model/test_data.rds")
```

# 3.0 Computing Correlation Matrix

Before loading the predictors into a predictive model, it is always a good practice to use correlation matrix to examine if there is sign of multicolinearity.

```{r}
mdata_nogeo <- mdata %>%
  st_drop_geometry()
corrplot::corrplot(cor(mdata_nogeo[, 2:17]), 
                   diag = FALSE, 
                   order = "AOE",
                   tl.pos = "td", 
                   tl.cex = 0.5, 
                   method = "number", 
                   type = "upper")
```

::: callout-note
The correlation matrix above shows that all the correlation values are below 0.8. Hence, there is no sign of multicolinearity.
:::

# 4.0 Retriving the Stored Data

```{r}
train_data <- read_rds("../../data/model/train_data.rds")
test_data <- read_rds("../../data/model/test_data.rds")
```

# 5.0 Building a non-spatial multiple linear regression

```{r}
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
summary(price_mlr)
```

```{r}
write_rds(price_mlr, "../../data/model/price_mlr.rds" ) 
```

# 6.0 gwr predictive method

In this section, you will learn how to calibrate a model to predict HDB resale price by using geographically weighted regression method of [**GWmodel**](https://cran.r-project.org/web/packages/GWmodel/) package.

## 6.1 Converting the sf data.frame to SpatialPointDataFrame

```{r}
train_data_sp <- as_Spatial(train_data)
train_data_sp
```

## 6.2 Computing adaptive bandwidth

Next, `bw.gwr()` of **GWmodel** package will be used to determine the optimal bandwidth to be used.

::: callout-note
The code chunk below is used to determine adaptive bandwidth and CV method is used to determine the optimal bandwidth.
:::

```{r}
#| eval: false
bw_adaptive <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=train_data_sp,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```

The result shows that 40 neighbour points will be the optimal bandwidth to be used if adaptive bandwidth is used for this data set.

```{r}
#| eval: false
write_rds(bw_adaptive, "../../data/model/bw_adaptive.rds")
```

## 6.3 Constructing the adaptive bandwidth gwr model

First, let us call the save bandwidth by using the code chunk below.

```{r}
bw_adaptive <- read_rds("../../data/model/bw_adaptive.rds")
```

Now, we can go ahead to calibrate the gwr-based hedonic pricing model by using adaptive bandwidth and Gaussian kernel as shown in the code chunk below.

```{r}
#| eval: false
gwr_adaptive <- gwr.basic(formula = resale_price ~
                            floor_area_sqm + storey_order +
                            remaining_lease_mths + PROX_CBD + 
                            PROX_ELDERLYCARE + PROX_HAWKER +
                            PROX_MRT + PROX_PARK + PROX_MALL + 
                            PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                            WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                            WITHIN_1KM_PRISCH,
                          data=train_data_sp,
                          bw=bw_adaptive, 
                          kernel = 'gaussian', 
                          adaptive=TRUE,
                          longlat = FALSE)
```

The code chunk below will be used to save the model in rds format for future use.

```{r}
#| eval: false
write_rds(gwr_adaptive, "../../data/model/gwr_adaptive.rds")
```

## 6.4 Retrieve gwr output object

The code chunk below will be used to retrieve the save gwr model object.

```{r}
gwr_adaptive <- read_rds("../../data/model/gwr_adaptive.rds")
```

The code below can be used to display the model output.

```{r}
gwr_adaptive
```

## 6.5 Converting the test data from sf data.frame to SpatialPointDataFrame

```{r}
test_data_sp <- test_data %>%
  as_Spatial()
test_data_sp
```

## 6.6 Computing adaptive bandwidth for the test data

```{r}
gwr_bw_test_adaptive <- bw.gwr(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                  data=test_data_sp,
                  approach="CV",
                  kernel="gaussian",
                  adaptive=TRUE,
                  longlat=FALSE)
```

## 6.7 Computing predicted values of the test data

```{r}
#| eval: false
gwr_pred <- gwr.predict(formula = resale_price ~
                          floor_area_sqm + storey_order +
                          remaining_lease_mths + PROX_CBD + 
                          PROX_ELDERLYCARE + PROX_HAWKER + 
                          PROX_MRT + PROX_PARK + PROX_MALL + 
                          PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                          WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
                          WITHIN_1KM_PRISCH, 
                        data=train_data_sp, 
                        predictdata = test_data_sp, 
                        bw=40, 
                        kernel = 'gaussian', 
                        adaptive=TRUE, 
                        longlat = FALSE)
```

# 7.0 Preparing coordinates data

## 7.1 Extracting coordinates data

The code chunk below extract the x,y coordinates of the full, training and test data sets.

```{r}
coords <- st_coordinates(mdata)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```

Before continue, we write all the output into rds for future used.

```{r}
#| eval: false
coords_train <- write_rds(coords_train, "../../data/model/coords_train.rds" )
coords_test <- write_rds(coords_test, "../../data/model/coords_test.rds" )
```

## 7.2 Droping geometry field

First, we will drop geometry column of the sf data.frame by using `st_drop_geometry()` of sf package.

```{r}
train_data <- train_data %>% 
  st_drop_geometry()
```

# 8.0 Calibrating Random Forest Model

In this section, you will learn how to calibrate a model to predict HDB resale price by using random forest function of [**ranger**](https://cran.r-project.org/web/packages/ranger/) package.

```{r}
#| eval: false
set.seed(1234)
rf <- ranger(resale_price ~ floor_area_sqm + storey_order + 
               remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE + 
               PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL + 
               PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
               WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
               WITHIN_1KM_PRISCH,
             data=train_data)
rf
```

```{r}
#| eval: false
write_rds(rf, "../../data/model/rf.rds")
```

```{r}
rf <- read_rds("../../data/model/rf.rds")
rf
```

# 9.0 Calibrating Geographical Random Forest Model

In this section, you will learn how to calibrate a model to predict HDB resale price by using `grf()` of [**SpatialML**](https://cran.r-project.org/web/packages/ranger/) package.

## 9.1 Calibrating using training data

The code chunk below calibrate a geographic ranform forest model by using `grf()` of **SpatialML** package.

```{r}
#| eval: false
set.seed(1234)
gwRF_adaptive <- grf(formula = resale_price ~ floor_area_sqm + storey_order +
                       remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE +
                       PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL +
                       PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                       WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                       WITHIN_1KM_PRISCH,
                     dframe=train_data, 
                     bw=55,
                     kernel="adaptive",
                     coords=coords_train)
```

Let’s save the model output by using the code chunk below.

```{r}
#| eval: false
write_rds(gwRF_adaptive, "../../data/model/gwRF_adaptive.rds")
```

The code chunk below can be used to retrieve the save model in future.

```{r}
#| eval: false
gwRF_adaptive <- read_rds("../../data/model/gwRF_adaptive.rds")
```

## 9.2 Predicting by using test data

### 9.2.1 Preparing the test data

The code chunk below will be used to combine the test data with its corresponding coordinates data.

```{r}
test_data <- cbind(test_data, coords_test) %>%
  st_drop_geometry()
```

### 9.2.2 Predicting with test data

Next, `predict.grf()` of spatialML package will be used to predict the resale value by using the test data and gwRF_adaptive model calibrated earlier.

```{r}
#| eval: false
gwRF_pred <- predict.grf(gwRF_adaptive, 
                           test_data, 
                           x.var.name="X",
                           y.var.name="Y", 
                           local.w=1,
                           global.w=0)
```

Before moving on, let us save the output into rds file for future use.

```{r}
#| eval: false
GRF_pred <- write_rds(gwRF_pred, "../../data/model/GRF_pred.rds")
```

### 9.2.3 Converting the predicting output into a data frame

The output of the `predict.grf()` is a vector of predicted values. It is wiser to convert it into a data frame for further visualisation and analysis.

```{r}
#| eval: false
GRF_pred <- read_rds("../../data/model/GRF_pred.rds")
GRF_pred_df <- as.data.frame(GRF_pred)
```

In the code chunk below, `cbind()` is used to append the predicted values onto test_data

```{r}
#| eval: false
test_data_p <- cbind(test_data, GRF_pred_df)
```

```{r}
#| eval: false
write_rds(test_data_p, "../../data/model/test_data_p.rds")
```

## 9.3 Calculating Root Mean Square Error

The root mean square error (RMSE) allows us to measure how far predicted values are from observed values in a regression analysis. In the code chunk below, rmse() of Metrics package is used to compute the RMSE.

```{r}
#| eval: false
rmse(test_data_p$resale_price, 
     test_data_p$GRF_pred)
```

## 9.4 Visualising the predicted values

Alternatively, scatterplot can be used to visualise the actual resale price and the predicted resale price by using the code chunk below.

```{r}
#| eval: false
ggplot(data = test_data_p,
       aes(x = GRF_pred,
           y = resale_price)) +
  geom_point()
```

:::callout-note
A better predictive model should have the scatter point close to the diagonal line. The scatter plot can be also used to detect if any outliers in the model.
:::